### 2016-03-25 hot_non-flaring_plasma_2/parameter_space_graphic.tex
outline changes, correcting heating stats section, adding figures, correcting figures

	\begin{tikzpicture}[node distance=2cm]
		%Draw the nodes
		\node (species) [ghost] {$\Sigma=\left\{
		\begin{array}{l}
			\mathrm{electron} \\
			\mathrm{ion} \\
			\mathrm{single}
		\end{array}
		\right.$};
		\node (ghost_ph) [ghost, below of=species] {};
		\node (alpha_pl) [ghost, left of=ghost_ph] {$\alpha=\left\{
		\begin{array}{l}
			-1.5 \\
			-2.0 \\
			-2.5
		\end{array}
		\right.$};
		\node (alpha_uni) [ghost, right of=ghost_ph] {uniform};
		\node (ghost_beta) [ghost, below of=alpha_pl]{};
		\node (beta_1) [ghost, right of=ghost_beta] {$\beta=1$};
		\node (beta_0) [ghost, left of=ghost_beta] {$\beta=0$};
		%Draw the arrows
		\draw [arrow] (species) -- (alpha_pl);
		\draw [arrow] (species) -- (alpha_uni);
		\draw [arrow] (alpha_pl) -- (beta_0);
		\draw [arrow] (alpha_pl) -- (beta_1);
		%\draw [arrow] (alpha_pl) -- (beta_2);	
	\end{tikzpicture}